### 2018-04-18 experiments.tex
more details on inference algs

The {\tt datarray} module \cite{datarray} is a prototype
implementation for numpy arrays with named axes to improve data
management, developed by the Berkeley Institute for Data Science.  As part of its code, it provides a set of algorithms for
inference in
Bayesian belief networks \cite{russell2016artificial}.  An earlier
version of these algorithms produced nondeterministic (and in some
cases incorrect) results due to dependence on the order of values in
an iterator over a Python dictionary (see Section \ref{sec:pnondet}).

Figure \ref{hashbug} shows TSTL code for generating inputs to the {\tt
  datarray} algorithms.  Running this harness using TSTL's {\tt
  --checkProcessDeterminism} flag required less than 10 seconds on
average to produce a test exhibiting process-level nondeterminism in
the {\tt calc\_marginals\_sumproduct} function (the only broken
algorithm).  Reducing this 60 step test to a minimal test of only 6 steps,
showing an extremely simple input producing the issue, 
required another 92 seconds.

>>>>

The {\tt datarray} module \cite{datarray} is a prototype
implementation for numpy arrays with named axes to improve data
management, developed by the Berkeley Institute for Data Science.  As part of its code, it provides a set of algorithms for
inference in
Bayesian belief networks \cite{russell2016artificial}.  An earlier
version of these algorithms produced nondeterministic (and in some
cases incorrect) results due to dependence on the order of values in
an iterator over a Python dictionary, on Python versions above 3.2, until 3.6 (see Section \ref{sec:pnondet}).

Figure \ref{hashbug} shows TSTL code for generating inputs to the {\tt
  datarray} algorithms.  The {\tt flatten_and_sort} function is needed because we care about actual differences in probability values, not simply the order of list, tuple, or dictionary items.  Running this harness using TSTL's {\tt
  --checkProcessDeterminism} flag required less than 10 seconds on
average to produce a test exhibiting process-level nondeterminism in
the {\tt calc\_marginals\_sumproduct} function (the only broken
algorithm).  Reducing this 60 step test to a minimal test of only 6 steps,
showing an extremely simple input producing the issue, 
required another 92 seconds.
